\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-08-05 analysis/paper/paper.Rmd

From the perspective of clinicians and their patients, the most relevant quantity is often the 5- or 10-year risk of having a certain event given the patient's particular circumstances, and not the hazard ratio between a treatment and control group. Therefore, to make sound clinical decisions, it is important to accurately estimate the *full* hazard function, which can then be used to estimate the cumulative incidence function (CIF). Since the parametric hazard is a smooth function of time, the CIF and the survival function estimates also vary smoothly over time.

@@

In this article, we present an \proglang{R} package that combines the ideas of Hanley & Miettinen into a simple interface. The purpose of the \pkg{casebase} package is to provide practitioners with an easy-to-use software tool to compute a patient's risk (or cumulative incidence) of an event, conditional on a particular patient's covariate profile. Our package retains the flexibility of case-base sampling and the familiar interface of the \code{glm} function. In addition, we provide extensive visualization tools.

>>>>

From the perspective of clinicians and their patients, the most relevant quantity is often the 5- or 10-year risk of having a certain event given the patient's particular circumstances, and not the hazard ratio between a treatment and control group. Therefore, to make sound clinical decisions, it is important to accurately estimate the *full* hazard function, which can then be used to estimate the cumulative incidence function (CIF). Using a parametric estimator of the hazard function leads to a smooth function of time; as a consequence, the CIF and the survival function estimates also vary smoothly over time.

@@

Using the framework of logistic regression for analysing time-to-event data opens the door to an extensive array of powerful modeling tools. Indeed, lasso and elastic-net regression can then be used to select variables that are associated with the hazard function. Logistic regression can also be replaced by multinomial regression to analyse data with multiple competing events.

In this article, we present an \proglang{R} package that combines the ideas of Hanley & Miettinen into a simple interface. The purpose of the \pkg{casebase} package is to provide practitioners with an easy-to-use software tool to compute a patient's risk (or cumulative incidence) of an event, conditional on a particular patient's covariate profile. Our package retains the flexibility of case-base sampling and the familiar interface of the \code{glm} function. It also provides tools for variable selection and competing-risk analysis. In addition, we provide extensive visualization tools.
